A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
Precisions about AST size
Correct numbers for parse evals

To roughly estimate the complexity of the code being evaluated,
Fig.~\ref{fig:ast-size} shows the distribution of the abstract syntax tree
size of the \c{expr} argument to \eval. The majority of inputs across core
and packages (\NbAstOnePercent of \eval calls and \NbAstOneCallSitePercent of call sites) have size 1. That is to say, the code being evaluated
consists of a single variable name or constant. A function call needs at
least three nodes and a list such as \c{c(1, 2, 3)} counts for 4 nodes. Inputs smaller than \AstSizeNineFive nodes account for 95\% of all
calls. Sizes in core are larger than in packages. It is due to core

@@

The pattern \c{eval(parse(...))} can be used for dynamic code loading. This
forms the core of \c{source} and \c{sys.source} functions in R that are commonly
used for loading code in R files in interactive settings. We investigated the
number of cases in which the output of \c{parse} and its variants is passed to
\eval, directly or transitively by tainting their output. We observed that
\PercentParsedEvals of the eval calls consume the result of calling \c{parse} on
a file and \PasteParsedEvals of the eval calls consume the result of calling
\c{parse} on a string. This corresponds to \PercentParsedCallSites and
\PasteParsedCallSite of the total \eval call sites. We also identified one
function in core R, \c{invokeRestartInteractively} that prompts the user for
input, parses it, and passes it to \eval.

>>>>

To roughly estimate the complexity of the code being evaluated,
Fig.~\ref{fig:ast-size} shows the distribution of the abstract syntax tree
size of the \c{expr} argument to \eval. The majority of inputs across core
and packages (\NbAstOnePercent of \eval calls and \NbAstOneCallSitePercent of call sites) have size 1. That is to say, the code being evaluated
consists of a single variable name or constant. A function call needs at
least three nodes and a list such as \c{c(1, 2, 3)} counts for 4 nodes. Inputs smaller than \AstSizeNineFive nodes account for 95\% of all
calls. Sizes in core are larger than in packages. Most large AST originate from library \emph{VGAM}, which is used for statistical modelling. It uses \eval to evaluate large matrices with values in a given environment.

@@

The pattern \c{eval(parse(...))} can be used for dynamic code loading. This
forms the core of \c{source} and \c{sys.source} functions in R that are commonly
used for loading code in R files in interactive settings. We investigated the
number of cases in which the output of \c{parse} and its variants is passed to
\eval, directly or transitively by tainting their output. This corresponds to \PercentParsedCallSites of the total \eval call sites and
\PercentParsedEvals of the eval calls. We observed that
very few of the eval call(\NbParseFilesRnd in total) consume the result of calling \c{parse} on
a file. Most of the eval calls consume the result of calling
\c{parse} on a string.  We also identified one
function in core R, \c{invokeRestartInteractively} that prompts the user for
input, parses it, and passes it to \eval.
